---
title: "360_giving_project"
author: 'a-abrhm'
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)
library(lubridate)
library(ggplot2)
library(RColorBrewer)
library(forcats)
library(gdata)
```

### Reading the data

Download the dataset from 360 giving's grantnav portal. Fundings made by Central,
Local and Devolved government types are taken for the period 2010 - 2022.
The downloaded dataset is fed into R.

```{r}
grant_data <- read.csv("grantnav-20230701154216.csv")
```

### Cleaning and preprocessing

Dropping insignificant columns

```{r}
grant_data <- select(grant_data, !c("Currency", "Amount.Applied.For", "Amount.Disbursed",
                                    "Recipient.Org.Charity.Number", "URL",
                                    "Planned.Dates.Start.Date", "Planned.Dates.End.Date",
                                    "Planned.Dates.Duration..months.","Actual.Dates.Start.Date" ,
                                     "Actual.Dates.End.Date", "Actual.Dates.Duration..months.",
                                    "Recipient.Org.Company.Number", "Recipient.Org.Postal.Code",
                                    "Recipient.Org.Location.0.Geographic.Code.Type",
                                    "Recipient.Org.Location.0.Geographic.Code",
                                    "Recipient.Org.Location.0.Name",
                                    "Recipient.Org.Location.1.Geographic.Code.Type",
                                    "Recipient.Org.Location.1.Geographic.Code",
                                    "Recipient.Org.Location.1.Name",                
                                    "Recipient.Org.Location.2.Geographic.Code.Type",
                                    "Recipient.Org.Location.2.Geographic.Code",
                                    "Recipient.Org.Location.2.Name",
                                    "Funding.Org.Postal.Code",
                                    "Grant.Programme.Code",
                                    "Grant.Programme.Title",
                                    "Grant.Programme.URL",
                                    "Beneficiary.Location.0.Name",
                                    "Beneficiary.Location.0.Country.Code",
                                    "Beneficiary.Location.0.Geographic.Code",
                                    "Beneficiary.Location.0.Geographic.Code.Type",                                                        "Beneficiary.Location.1.Name",             
                                    "Beneficiary.Location.1.Country.Code",                          
                                    "Beneficiary.Location.1.Geographic.Code",
                                    "Beneficiary.Location.1.Geographic.Code.Type",
                                    "Beneficiary.Location.2.Name",
                                    "Beneficiary.Location.2.Country.Code",
                                    "Beneficiary.Location.2.Geographic.Code",
                                    "Beneficiary.Location.2.Geographic.Code.Type",
                                    "Beneficiary.Location.3.Name",
                                    "Beneficiary.Location.3.Country.Code",
                                    "Beneficiary.Location.3.Geographic.Code",
                                    "Beneficiary.Location.3.Geographic.Code.Type",
                                    "Beneficiary.Location.4.Name",
                                    "Beneficiary.Location.4.Country.Code",
                                    "Beneficiary.Location.4.Geographic.Code",
                                    "Beneficiary.Location.4.Geographic.Code.Type",
                                    "Beneficiary.Location.5.Name",                     
                                    "Beneficiary.Location.5.Country.Code",
                                    "Beneficiary.Location.5.Geographic.Code",
                                    "Beneficiary.Location.5.Geographic.Code.Type",
                                    "Beneficiary.Location.6.Name",
                                    "Beneficiary.Location.6.Country.Code",
                                    "Beneficiary.Location.6.Geographic.Code",
                                    "Beneficiary.Location.6.Geographic.Code.Type",
                                    "Beneficiary.Location.7.Name",
                                    "Beneficiary.Location.7.Country.Code",
                                    "Beneficiary.Location.7.Geographic.Code",
                                    "Beneficiary.Location.7.Geographic.Code.Type",
                                    "From.An.Open.Call.", "Data.Source",  "The.following.fields.are.not.in.the.360.Giving.Standard.and.are.added.by.GrantNav.",
                                    "Publisher.Name", "Recipient.Region",
                                    "Recipient.District",                      
                                    "Recipient.District.Geographic.Code",
                                    "Recipient.Ward",
                                    "Recipient.Ward.Geographic.Code",
                                    "Retrieved.for.use.in.GrantNav",
                                    "Funding.Org..Org.Type",
                                    "Funding.Org..Canonical.Org.ID",
                                    "Funding.Org..Canonical.Name",
                                    "Recipient.Org..Date.Registered",
                                    "Recipient.Org..Date.Removed",
                                    "Recipient.Org..Org.ID.s.",
                                    "Recipient.Org..Latest.Income",
                                    "Recipient.Org..Latest.Income.Date",
                                    "Recipient.Org..Org.Type",
                                    "Recipient.Org..Canonical.Org.ID",
                                    "Recipient.Org..Canonical.Name",
                                    "Recipient.Org..Registered.Postcode",
                                    "Recipient.Org..Data.Source","License..see.note.",
"Note..this.file.also.contains.OS.data...Crown.copyright.and.database.right.2016..Royal.Mail.data...Royal.Mail.copyright.and.Database.right.2016..National.Statistics.data...Crown.copyright.and.database.right.2015...2016..see.http...grantnav.threesixtygiving.org.datasets..for.more.information."
                                    ))
```

The recipient.org.name column contains duplicate names with different
cases. They also have 2 types of recipient.org.identifiers (one starting
with 360G-... and the other starting with GB-...). Since all these point
to the same recipient organisation, we are making them have the same
identifier (one starting with GB-...)

```{r}
grant_data$Recipient.Org.Name <- tolower(grant_data$Recipient.Org.Name)
```

The recipient.org.name column has entries [placeholder] and [foreign
affairs]. Replacing those values with the corresponding
recipient.org.identifier

```{r}
grant_data <- grant_data %>% mutate(Recipient.Org.Name =
                                      ifelse(str_detect(Recipient.Org.Name,
                                                        "^\\[(.*)\\]$"),
                                             Recipient.Org.Identifier,
                                             Recipient.Org.Name))
```

The university recipients have multiple names and identifiers, all of
which refers to the same entity. Therefore making all such rows of the
same entity have the same organisation name and identifier.

```{r}
grant_data <- grant_data %>% mutate(Recipient.Org.Name.Temp = str_extract(Recipient.Org.Name,
                                         "^(the u|u)niversity of \\w+"))
grant_data <- grant_data %>%
  mutate(Recipient.Org.Name = ifelse(!is.na(Recipient.Org.Name.Temp),
                                     str_extract(Recipient.Org.Name.Temp,
                                                 "university of \\w+"),
                                     Recipient.Org.Name)) %>% select(!(Recipient.Org.Name.Temp))
# grouping by identifier and making the organisation name same
recipient_names_and_identifiers <- grant_data %>% count(Recipient.Org.Name,
                                                  Recipient.Org.Identifier)
for(i in 1:nrow(recipient_names_and_identifiers)) {
  if (i > 1 && recipient_names_and_identifiers$Recipient.Org.Name[i] ==
      recipient_names_and_identifiers$Recipient.Org.Name[i - 1]) {
    next
  }
  identifiers <- recipient_names_and_identifiers$Recipient.Org.Identifier[
    recipient_names_and_identifiers$Recipient.Org.Name ==
      recipient_names_and_identifiers$Recipient.Org.Name[i]
  ]
  grant_data$Recipient.Org.Identifier[grant_data$Recipient.Org.Name ==
                    recipient_names_and_identifiers$Recipient.Org.Name[i]] <-
    identifiers[length(identifiers)]
}
```

Converting columns to their correct datatypes

```{r warning=FALSE}
grant_data <- mutate(grant_data, Amount.Awarded = as.double(Amount.Awarded)) %>% na.omit()
```

Dropping rows with "Amount.Awarded" less than or equal to zero ( new
case where funds are one digit. So removing funds less than 100 pounds )

```{r}
grant_data <- filter(grant_data, Amount.Awarded > 100)
```

Changing the "Award.Date" column to datetime. The original data set has
dates in 2 formats (d-m-y and m-d-y). So the following logic is used to
parse dates in any order of day, month and year to date.

Function that converts date_strings in any order to date

```{r}
convert_date_strings <- function(date_strings, separator = "\\/|\\-") {
  ds <- as.vector(sapply(str_split(date_strings, separator), as.integer))
  # max value day can take is 31
  # max value month can take is 12
  # so anything greater than 31 is taken as year
  # Also we assume by that the month value will come in the middle
  y <- ds[ds > 31]
  m <- ds[ds < 32][rep(c(FALSE, TRUE), length(date_strings))]
  d <- ds[ds < 32][rep(c(TRUE, FALSE), length(date_strings))]
  mismatched_indices <- which(m > 12)
  if (length(mismatched_indices)) {
    m[mismatched_indices] <- ds[ds < 32][rep(c(TRUE, FALSE),
                                             length(date_strings))][mismatched_indices]
    d[mismatched_indices] <- ds[ds < 32][rep(c(FALSE, TRUE),
                                             length(date_strings))][mismatched_indices]
  }
  return(make_date(y, m, d))
}
```

```{r converting date string to date}
grant_data <- mutate(grant_data, Award.Date = convert_date_strings(Award.Date))
```

Creating a column to store the year each grant was made

```{r}
grant_data <- mutate(grant_data, "Award.Year" = year(Award.Date))
```

Removing grants made before 2010

```{r}
grant_data <- filter(grant_data, Award.Year >= 2010)
```

### Sectors

Classifying grants made to different sectors or themes

Creating a dictionary of sector and keywords of that sector

Not including "environment" as the keywords have overlap Have to
identify which sector to assign if there is overlap with multiple
categories

```{r sector-keywords dictionary}
sector_keywords_dict <- list(
  "education" = c("education", "school", "university", "research"),
  "covid" = c("covid"),
  "homeless/housing" = c("house", "housing", "home", "tenant", "rough sleeper"),
  "business" = c("business", "trade", "market"),
  "artistic_cause" = c("art", "craft"),
  "animal_welfare_and_disease" = c("animal", "fauna", "zoo", "wildlife"),
  "blindness" = c("blind", "visually impaired", "eye"),
  "care_of_elderly" = c("elderly", "senior citizen"),
  "children" = c("child", "young people"),
  "community_development" = c("community development", "community"),
  "conservation/preservation" = c("conservation", "perservation"),
  "deafness" = c("deaf"),
  "destitution" = c("destitute", "destitution", "poverty", "asylum"),
  "refugees" = c("refugee"),
  "migrant" = c("migrant"),
  "disables/handicapped" = c("disability", "disabilities", "handicap", "mobility",
                             "lonely", "health problem"),
  "emergency_appeal" = c("emergency appeal"),
  "employment_and_job_creation" = c("employment", "job"),
  "human_rights" = c("human right"),
  "medical_care_and_research" = c("medical", "medicine", "hospital", "sick", "health",
                                  "doctor", "NHS", "illness", "disease"),
  "mental_health" = c("mental health"),
  "violence_abuse_and_crime_prevention" = c("violence", "abuse", "crime", "victim",
                                            "prison", "sexual abuse", "rape"),
  "water_sanitation_and_hygiene" = c("water", "sanitation", "hygiene"),
  "transportation_and_logistics" = c("transport", "logistic", "freight", "biofuel",
                                     "mini bus", "SEFT", "supply chain", "highways",
                                     "bike", "LSTF", "crossrail", "TFL", "road safety",
                                     "transit", "vechicle", "truck", "CAA", "van",
                                     "bus", "energy saving"),
  "agriculture" = c("agricultur", "horticulture", "cultivat"),
  "culture" = c("tradition", "culture", "heritage", "history"),
  "electric_vechicles" = c("home chargers", "electric vehicle", "EV", "ULEV", 
                           "plugged in"),
  "sports" = c("play", "activity"),
  "awareness campaign" = c("awareness", "campaign"),
  "festival" = c("festival")
)
```

Function that returns possible sectors for a grant based on the title
and description

```{r function to filter by keyword match}
find_sectors <- function(title = "", description = "") {
  sectors <- sapply(sector_keywords_dict, function(item) {
    str_detect(title, paste("(?i)(", paste0(item, collapse = "|"), ")", sep = "")) |
      str_detect(description, paste("(?i)(", paste0(item, collapse = "|"), ")", sep = ""))
  })
  return(paste0(names(sectors[sectors == TRUE]), collapse = ", "))
}
```

Creating a new column that will store possible sectors a grant is made
to.

```{r}
grant_data$Possible.Sectors <- ""
for (i in 1:nrow(grant_data)) {
  grant_data$Possible.Sectors[i] <- find_sectors(grant_data$Title[i],
                                                 grant_data$Description[i])
}
```

Converting the rows with multiple sectors and containing covid to only
covid.

```{r}
grant_data$Possible.Sectors[str_detect(grant_data$Possible.Sectors, "covid")] <- "covid"
```

The sector classification function assigns multiple categories for some
rows. Such rows must be labelled correctly by manual supervision. For now,
only taking the first sector in case of multiple sectors.

Note:- The order of sectors in the sector_keywords_dict is important as we are
        assigning the first matching sector. The order should be carefully determined
        if you see there is overlapping of keywords in different categories.

```{r}
grant_data$Sector <- sapply(str_split(grant_data$Possible.Sectors, ","), "[[", 1)
```

### Visualizations

#### Funder stats

Grouping the data by grants made by each funder

```{r}
funder_stats <- grant_data %>% group_by(Funding.Org.Identifier) %>%
  summarise(Funding.Org.Name = first(Funding.Org.Name),
            Total.Amount = sum(Amount.Awarded) / 1000000000,
            Total.Grants = n()/1000)
```

```{r}
# selecting the funder names from the sorted dataframe
ordered_funders <- arrange(funder_stats, Total.Amount) %>% select(Funding.Org.Name)
# making organisation name as a factor and
# ordering the factors by the levels described by ordered_funders (and not alphabetically)
funder_stats$Funding.Org.Name <- factor(funder_stats$Funding.Org.Name,
                                        levels = ordered_funders$Funding.Org.Name)
# ordering the funder stats by organisation name factor levels so that when plotted,
# the organisations will not be ordered alphabetically on the axis
funder_stats <- arrange(funder_stats, Funding.Org.Name)

# line graph
ggplot(funder_stats, aes(x = Total.Amount, y = Funding.Org.Name, group = 1)) +
  geom_line(orientation = "y", color = "red") + geom_point(color = "blue") +
  xlim(0,50) + labs(x = "Total amount (in billions)", y = "Funding orgainsations")
# bar graph
ggplot(funder_stats, aes(y = Funding.Org.Name, x = Total.Amount)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_x_continuous(n.breaks = 10) + 
  labs(y = "Funding orgainsations", x = "Total amount (in billions)")

# ordering funder stats in decreasing order of total grants
ordered_funders <- arrange(funder_stats, Total.Grants) %>% select(Funding.Org.Name)
# changing the order of levels in organisation name factor
funder_stats$Funding.Org.Name <- reorder.factor(funder_stats$Funding.Org.Name,
                                               new.order = ordered_funders$Funding.Org.Name)
# ordering by factor levels
funder_stats <- arrange(funder_stats, Funding.Org.Name)

# line graph
ggplot(funder_stats, aes(x = Total.Grants, y = Funding.Org.Name, group = 1)) +
  geom_line(orientation = "y", color = "red") + geom_point(color = "blue") +
  xlim(0,60) + labs(x = "Total grants made", y = "Funding orgainsations")

# bar graph
ggplot(funder_stats, aes(y = Funding.Org.Name, x = Total.Grants)) +
  geom_bar(stat = "identity", fill = "yellowgreen") +
  scale_x_continuous(n.breaks = 10) +
  scale_y_discrete(labels = funder_stats$Funding.Org.Name) +
  labs(y = "Funding orgainsations", x = "Total Grants made")
```

#### Yearly statistics

```{r}
yearly_stats <- grant_data %>% group_by(Award.Year) %>%
  summarise(Avg.Amount.Awarded = mean(Amount.Awarded) / 1000000,
            Total.Amount.Awarded = sum(Amount.Awarded) / 1000000000,
            Total.Grants.Given = n() / 1000)

ggplot(yearly_stats, aes(x = Award.Year, y = Avg.Amount.Awarded)) +
  geom_line(color = "yellowgreen") + geom_point(color = "blue") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Average Amount Awarded (in millions)")

ggplot(yearly_stats, aes(x = Award.Year, y = Total.Amount.Awarded)) +
  geom_line(color = "yellowgreen") + geom_point(color = "blue") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Total Amount Awarded (in billions)")

ggplot(yearly_stats, aes(x = Award.Year, y = Total.Grants.Given)) +
  geom_line(color = "yellowgreen") + geom_point(color = "blue") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Total Grants Given (in thousands)")
```

```{r}
ggplot(yearly_stats, aes(x = Award.Year, y = Avg.Amount.Awarded)) +
   geom_bar(stat = "identity", fill = "yellowgreen") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Average Amount Awarded (in millions)")
  
ggplot(yearly_stats, aes(x = Award.Year, y = Total.Amount.Awarded)) +
  geom_bar(stat = "identity", fill = "yellowgreen") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Total Amount Awarded (in billions)")

ggplot(yearly_stats, aes(x = Award.Year, y = Total.Grants.Given)) +
  geom_bar(stat = "identity", fill = "yellowgreen") +
  scale_x_continuous(n.breaks = nrow(yearly_stats)) +
  labs(x = "Award Year", y = "Total Grants Given (in thousands)")
```

#### Sector stats

```{r}
sector_stats <- grant_data %>% group_by(Sector) %>%
  summarise(Total.Amount.Awarded = sum(Amount.Awarded) / 1000000000,
            Total.Grants.Given = n() / 1000) %>% arrange(desc(Total.Amount.Awarded))
sector_stats <- filter(sector_stats, Sector != "")

ordered_sectors <- arrange(sector_stats, desc(Total.Amount.Awarded)) %>%
  select(Sector)
# making organisation name as a factor and
# ordering the factors by the levels described by ordered_sectors (and not alphabetically)
sector_stats$Sector <- factor(sector_stats$Sector,
                                        levels = ordered_sectors$Sector)
# ordering the sector stats by organisation name factor levels so that when plotted,
# the organisations will not be ordered alphabetically on the axis
sector_stats <- arrange(sector_stats, Sector)

ggplot(sector_stats, aes(x = Sector, y = Total.Amount.Awarded)) +
   geom_bar(stat = "identity", fill = "yellowgreen") +
  guides(x = guide_axis(angle = 60)) +
  labs(x = "Sector", y = "Total Amount Awarded (in billions)")

# ordering sector stats in decreasing order of total grants
ordered_sectors <- arrange(sector_stats, desc(Total.Grants.Given)) %>%
  select(Sector)
# changing the order of levels in organisation name factor
sector_stats$Sector <- reorder.factor(sector_stats$Sector,
                                               new.order = ordered_sectors$Sector)
# ordering by factor levels
sector_stats <- arrange(sector_stats, Sector)
  
ggplot(sector_stats, aes(x = Sector, y = Total.Grants.Given)) +
  geom_bar(stat = "identity", fill = "yellowgreen") +
  guides(x = guide_axis(angle = 60)) +
  labs(x = "Sector", y = "Total Grants Given (in thousands)")
```

Education sector has received the most grants and total amount, followed
by artistic cause and then covid.

#### Recipient statistics

```{r}
recipient_stats_decomposed <- grant_data %>% group_by(Recipient.Org.Identifier,
                                                      Funding.Org.Identifier) %>%
  summarise(Recipient.Org.Name = first(Recipient.Org.Name),
            Amount.Received = sum(Amount.Awarded) / 1000000000,
            Grants.Received = n())

recipient_stats <- recipient_stats_decomposed %>% group_by(Recipient.Org.Identifier) %>%
  summarise(Recipient.Org.Name = first(Recipient.Org.Name),
            Total.Amount.Received = sum(Amount.Received),
            Total.Grants.Received = sum(Grants.Received),
            Total.Funders = n())

# ordering by amount received

ordered_recipients <- arrange(recipient_stats, desc(Total.Amount.Received)) %>%
  select(Recipient.Org.Identifier)
recipient_stats$Recipient.Org.Identifier <- factor(
  recipient_stats$Recipient.Org.Identifier,
  levels = ordered_recipients$Recipient.Org.Identifier)
recipient_stats <- arrange(recipient_stats, Recipient.Org.Identifier)

ggplot(recipient_stats[1:10,], aes(y = Recipient.Org.Identifier,
                                   x = Total.Amount.Received)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_y_discrete(labels = recipient_stats$Recipient.Org.Name) +
  labs(y = "Recipient orgainsations", x = "Total amount (in billions)")

recipient_stats_decomposed$Recipient.Org.Identifier <-
  factor(recipient_stats_decomposed$Recipient.Org.Identifier,
         levels = ordered_recipients$Recipient.Org.Identifier)

recipient_stats_decomposed <- arrange(recipient_stats_decomposed,
                                      Recipient.Org.Identifier,
                                      desc(Amount.Received))

ggplot(recipient_stats_decomposed[1:52,], aes(y = Recipient.Org.Identifier,
                                   x = Amount.Received,
                                   fill = Funding.Org.Identifier)) +
  geom_bar(stat = "identity") +
  scale_y_discrete(labels = recipient_stats$Recipient.Org.Name) +
  labs(y = "Recipient orgainsations", x = "Total amount (in billions)")

# ordering by grants received

ordered_recipients <- arrange(recipient_stats, desc(Total.Grants.Received)) %>%
  select(Recipient.Org.Identifier)
recipient_stats$Recipient.Org.Identifier <- reorder.factor(
  recipient_stats$Recipient.Org.Identifier,
  new.order = ordered_recipients$Recipient.Org.Identifier)
recipient_stats <- arrange(recipient_stats, Recipient.Org.Identifier)

ggplot(recipient_stats[1:10,], aes(y = Recipient.Org.Identifier,
                                   x = Total.Grants.Received)) +
  geom_bar(stat = "identity", fill = "darkgreen") +
  scale_y_discrete(labels = recipient_stats$Recipient.Org.Name) +
  labs(y = "Recipient orgainsations", x = "Total grants")


recipient_stats_decomposed$Recipient.Org.Identifier <-
  reorder.factor(recipient_stats_decomposed$Recipient.Org.Identifier,
         new.order = ordered_recipients$Recipient.Org.Identifier)

recipient_stats_decomposed <- arrange(recipient_stats_decomposed,
                                      Recipient.Org.Identifier,
                                      desc(Grants.Received))

# we are only taking the top 10 recipients (it corresponds to the 1st 52 rows)
recipient_stats_decomposed <- recipient_stats_decomposed[1:52,]
top_funders <- recipient_stats_decomposed %>%
  group_by(Funding.Org.Identifier) %>% summarise(Total.Grants =
                                                   sum(Grants.Received)) %>%
  arrange(Total.Grants)
top_funders$Funding.Org.Name <- ""
#for (i in 1:nrow(top_funders)) {
#  top_funders$Funding.Org.Name[i] <- funder_stats$Funding.Org.Name[
#    funder_stats$Funding.Org.Identifier == top_funders$Funding.Org.Identifier[i]
#  ]
#}
top_funders <- top_funders$Funding.Org.Identifier
recipient_stats_decomposed$Funding.Org.Identifier <- factor(
  recipient_stats_decomposed$Funding.Org.Identifier, levels = top_funders
)

ggplot(recipient_stats_decomposed, aes(y = Recipient.Org.Identifier,
                                   x = Grants.Received,
                                   fill = Funding.Org.Identifier)) +
  geom_bar(stat = "identity") +
  scale_y_discrete(labels = recipient_stats$Recipient.Org.Name) +
  scale_fill_manual(values = c("#de2d26", "#756bb1", "#e6550d", "#2c7fb8", "#31a354",
                               "#c51b8a", "#999999", "#E69F00", "#fec44f", "#7fcdbb",
                               "#99d8c9", "#fa9fb5", "#a6bddb", "#0d0f66")) +
  labs(y = "Recipient orgainsations", x = "Total grants")
```

Calculating the sector statistics for each year so as to get data for time series
prediction. The resulting data frame will consists of the total amount and total
grants received by each sector in each year. It should be noted that some sectors
did not receive any funding in some of the years.

```{r}
yearly_sector_stats <- grant_data %>% group_by(Sector, Award.Year) %>%
  summarise(Total.Amount.Awarded = sum(Amount.Awarded) / 1000000000,
            Total.Grants.Given = n() / 1000) %>% filter(Sector != "")
```

Writing the data frames into csv files for future reference and use.

```{r}
write.csv(grant_data, "grant_data.csv")
write.csv(funder_stats, "funder_stats.csv")
write.csv(yearly_stats, "yearly_stats.csv")
write.csv(sector_stats, "sector_stats.csv")
write.csv(recipient_stats_decomposed, "recipient_stats_decomposed.csv")
write.csv(recipient_stats, "recipient_stats.csv")
write.csv(yearly_sector_stats, "yearly_sector_stats.csv")
```

